---
Author: "Kathleen Moriarty"
Course: "Real-World Data (395-0102-00L) Autumn 2022"
Location: "D-HEST, ETH Zurich"
Date: "29 November 2022"
Title: "Explore CMS Synthetic Claims Data: Optional Exercises"
Desc: "Explore the data and code independently, trying to have an initial answer to the exercise questions."
output: html_notebook
---

Homework Notes

All of these solutions are oversimplification of the responses to the research questions at hand.
Thorough cleaning/filter/imputing of data would be needed as well as literature review of what is already known would be needed before drawing any conclusions.


1. Create other plots of interest for the exploratory section. (medium)

2. Check for missingness of another variable of interest (easy). 
Without pursuing a deeper analysis, in your opinion, does this variable appear to be missing at random or not?
See section “# Missingness” from the 2_explore_data.Rmd file

```{r}
# Check missingness of admiting ICD9 diagnosis code.
# Missing_compare: Comparisons for continuous data use a Kruskal Wallis and for discrete data a chi-squared test.
outcome <- 'ICD9_DGNS_CD_1'
explanatory <- c('female', 'race', 'age', 'continuous_coverage','SP_ALZHDMTA', 'SP_DEPRESSN')
op_claims %>% 
  left_join(members, by = c("ID", "YR")) %>% 
  missing_compare(outcome, explanatory)

```

3. Show mean claim amount for diagnoses related to diabetes (difficult)
In file 4_prepare_cohort_analysis.Rmd, refer to lines 129-133 (with comment “# Find any claims for diabetes”) to find the diabetes ICD9 diagnosis codes 
Refer to 2_explore_data.Rmd for exploration code with claims. 
Optional: include the carrier claims in this as well (commented out parts)

```{r}

# Find any claims for diabetes
dia_icd9 <- c('249', '250', '3572', '36201', '36202', '36641')

# Select the inpatient/outpatient claims with these codes
dia_claims <- ip_op_claims %>% 
  filter(if_any(starts_with("ICD9_DGNS"), ~ . %in% dia_icd9)) %>% 
   mutate(comorbid_type = "dia")

# Select the mean claim amount
dia_claims %>% 
  group_by(claim_type) %>% 
  summarize(mean_medicare_amt = mean(medicare_amt),
            mean_payer_amt = mean(payer_amt),
            mean_patient_amt = mean(patient_amt))

# Interpretation (oversimplification)
# The results show that mean medicare, payer, patient amount per claim for all claims related to diabetes.
# We can compare this with another chronic condition

# Find any claims for hypertension
hyp_claims1 <- ip_op_claims %>% 
  filter(if_any(starts_with("ICD9_DGNS"), ~ str_starts(., pattern = "4011")))
hyp_claims2 <- ip_op_claims %>% 
  filter(if_any(starts_with("ICD9_DGNS"), ~ str_starts(., pattern = "4019")))
hyp_claims3 <- ip_op_claims %>% 
  filter(if_any(starts_with("ICD9_DGNS"), ~ str_starts(., pattern = "4010")))
hyp_claims <- bind_rows(hyp_claims1, hyp_claims2, hyp_claims3) %>% 
  mutate(comorbid_type = "hyp")
          
# Select the mean claim amount
hyp_claims %>% 
  group_by(claim_type) %>% 
  summarize(mean_medicare_amt = mean(medicare_amt),
            mean_payer_amt = mean(payer_amt),
            mean_patient_amt = mean(patient_amt))

# Why would there be a difference with hyptension

```

4. Plot the top 10 diagnosis codes by cost burden  (difficult)
Focus only on inpatient & outpatient claims
Group the dataframe by diagnosis code and summarize the claim amounts
Plot a bar graph of the 10
Do the results surprise you? What could explain these results?
Explore the data on your own and find something of interest to you (difficult)

```{r}

ip_op_claims %>% 
  group_by(ADMTNG_ICD9_DGNS_CD) %>% 
  summarize(mean_all_cost = mean(medicare_amt + payer_amt + patient_amt)) %>% 
  arrange(desc(mean_all_cost)) %>% 
  slice_head(n=10)
  
# This code would also work 
ip_op_claims %>% 
  group_by(ADMTNG_ICD9_DGNS_CD) %>% 
  summarize(mean_all_cost = mean(medicare_amt + payer_amt + patient_amt)) %>% 
  slice_max(mean_all_cost, n=10)

44282: Aneurysm of subclavian artery
801.25: Closed fracture of base of skull with subarachnoid, subdural, and extradural hemorrhage, with prolonged [more than 24 hours] loss of consciousness, without return to pre-existing conscious level
94520: Blisters, epidermal loss [second degree] of lower limb [leg], unspecified site
# plot
  
```

Add a new chunk by clicking the *Insert Chunk* button on the toolbar or by pressing *Ctrl+Alt+I*.

When you save the notebook, an HTML file containing the code and output will be saved alongside it (click the *Preview* button or press *Ctrl+Shift+K* to preview the HTML file).

The preview shows you a rendered HTML copy of the contents of the editor. Consequently, unlike *Knit*, *Preview* does not run any R code chunks. Instead, the output of the chunk when it was last run in the editor is displayed.
