---
Author: "Kathleen Moriarty"
Course: "Introducation to Data Science and RWD"
Location: "Roche Rotkreuz"
Date: "30 March 2023"
Desc: 'Clean datasets: make new variables, change variable types'
Title: Explore CMS Synthetic Claims
output: html_document:
---

# Missingness
### Great Explanations and Vignettes: https://cran.r-project.org/web/packages/finalfit/vignettes/missing.html

```{r}

members %>% 
  ff_glimpse()

ip_claims %>% 
  ff_glimpse()
# We would need to dig deeper into these 68 claims that are missing vital information.

ip_claims %>% 
  missing_plot()

ip_claims %>% 
  select(-starts_with("HCPCS_CD")) %>% 
  missing_plot()

# Join ip_claims with demographic information to see if the missing variables are at random when looking at demographics
ip_claims %>% 
  left_join(members, by = c("ID", "YR")) %>% 
  select(-starts_with("HCPCS_CD")) %>% 
  missing_plot()

# Check missingness of admiting ICD9 diagnosis code.
# Missing_compare: Comparisons for continuous data use a Kruskal Wallis and for discrete data a chi-squared test.
outcome <- 'ADMTNG_ICD9_DGNS_CD'
explanatory <- c('female', 'race', 'age', 'continuous_coverage','SP_ALZHDMTA')
ip_claims %>% 
  left_join(members, by = c("ID", "YR")) %>% 
  missing_compare(outcome, explanatory)

# Check other variables of interest
  
```

# Patient Demographics + Disease History


## Gender by Chronic Condition 


```{r}

# Make the data 'long' rather than 'wide'
# i.e. Convert the member table from one row per member per year to one row per member per chronic condition
# We want to add the number of chronic conditions so 1 indicates that it exists, otherwise 0.
members_long <- members %>%
  pivot_longer(cols = starts_with("SP"),
               names_to = "chronic_cond", 
               values_to = "chronic_cond_ind") %>% 
  mutate(cc_ind = if_else(chronic_cond_ind == 1, 1, 0)) #chronic_cond_ind: 2 means No, 1 means Yes

# Summarize the data by each group (chronic condition, gender, year)
# Calculate the proportion of the records in those groups
members_long_sum <- members_long %>% 
  group_by(chronic_cond, female, YR) %>% 
  summarize(freq = sum(cc_ind),
            tot = n(),
            prop = freq/n()) %>%
  mutate(freq_plot = case_when(
    female == "0" ~ -freq,
    TRUE ~ freq
  )) %>% 
  ungroup()

# Plot the results
ggplot(members_long_sum, aes(x=chronic_cond, y=prop, fill = female)) +
  geom_col(position = "dodge") + 
  theme_bw() +
  theme(legend.title=element_blank(), 
        axis.text.x = element_text(angle = 90, vjust = 0.5, hjust=1)) +
  scale_fill_discrete(labels = c("Male", "Female")) +
  labs(x='Chronic Conditions', y='', title='Chronic Conditions by Gender and Year') +
  facet_wrap(~YR)

```

### What questions arise from this plot?

## Age by Gender

```{r}
ggplot(members, aes(x=age, fill = female)) +
  geom_histogram(alpha = 0.7) +
  theme_bw() +
  labs(x='Age', y='', title='Age by Gender and Year') +
  scale_fill_discrete(labels = c("Male", "Female")) +
  facet_wrap(~YR)

```

### What questions arise from this plot?

# Claim distribution

## Claim Distribution by Claim Type

```{r}

ip_op_claims %>% 
  filter(YR == 2008) %>% 
  group_by(claim_type) %>% 
  summarize(tot_patient_amt = sum(patient_amt)) %>% 
ggplot(., aes(x=claim_type, y=tot_patient_amt, fill = claim_type)) +
  geom_col(position = "dodge") + 
  theme_bw() +
  theme(legend.title=element_blank()) +
  scale_fill_discrete(labels = c("Inpatient", "Outpatient")) +
  labs(x='Claim Type', y='', title='Total Claim Cost by Claim Type')

  
ip_op_claims %>% 
  filter(YR == 2008) %>% 
  group_by(claim_type) %>% 
  summarize(tot_num_claims = n()) %>% 
ggplot(., aes(x=claim_type, y=tot_num_claims, fill = claim_type)) +
  geom_col(position = "dodge") + 
  theme_bw() +
  theme(legend.title=element_blank()) +
  scale_fill_discrete(labels = c("Inpatient", "Outpatient")) +
  labs(x='Claim Type', y='', title='Total Number of Claims by Claim Type')

```

## Other Plots of Interest

## Questions: How could you display 
## 1. Claim Amount by Selected Diagnosis Codes
## 2. Top 10 diagnoses by cost burdern
## 3. What other relationships would be interesting?

