---
Author: "Kathleen Moriarty"
Course: "Introducation to Data Science and RWD"
Location: "Roche Rotkreuz"
Date: "30 March 2023"
Title: "Explore CMS Synthetic Claims Data: Prepare Cohort for Analysis"
Desc: "Prepare the Cohort for analysis: combine cohort with demographics and  select all claims records 12+/- index date"
output: html_notebook
---


# Select all IP, OP, Carrier claims for that cohort that would be comorbidities & post procedure claims

```{r}

# Join cohort table with claims table and select only claims wihtin +/- 12 months of index date
cohort_claims <- cohort %>% 
  #left_join(ip_op_carrier_claims, by = c("ID")) %>% 
  left_join(ip_op_claims, by = c("ID")) %>% 
  mutate(year_plus_ind = index_dt %m+% months(12),
         year_minus_ind = index_dt %m-% months(12))

cohort_comorbidities <- cohort_claims %>% 
  filter(CLM_FROM_DT >= year_minus_ind &
           CLM_FROM_DT < index_dt)

cohort_post_procedure <- cohort_claims %>% 
  filter(CLM_FROM_DT <= year_plus_ind &
           CLM_FROM_DT > index_dt)


```

# Filter the claims for outcome cardiac event diagnoses
### angina, heart failure, ischemic heart disease, or cardiac arrest

```{r}

# Find any claims for angina
ang_icd9 <- c('413')
ang_claims <- cohort_post_procedure %>% 
  filter(if_any(starts_with("ICD9_DGNS"), ~ . %in% ang_icd9)) %>% 
  mutate(outcome_type = "ang")

# Find any claims for heart failure
hf_icd9 <- c('39891', '40201', '40211', '40291',
              '40401', '40411', '40491', 
              '40403', '40413', '40493')
hf_claims1 <- cohort_post_procedure %>% 
  filter(if_any(starts_with("ICD9_DGNS"), ~ . %in% hf_icd9))
hf_claims2 <- cohort_post_procedure %>% 
  filter(if_any(starts_with("ICD9_DGNS"), ~ str_starts(., pattern = "428")))

hf_claims <- bind_rows(hf_claims1, hf_claims2) %>% 
  mutate(outcome_type = "hf")

# Find any claims for ihd
is_heart_icd9_dgns <- c('41000', '41001', '41002', 
              '41010', '41011', '41012',
              '41020', '41021', '41022', 
              '41030', '41031', '41032', 
              '41040', '41041', '41042', 
              '41050', '41051', '41052', 
              '41060', '41061', '41062', 
              '41070', '41071', '41072', 
              '41080', '41081', '41082', 
              '41090', '41091', '41092', 
              '4110', '4111', '41181', '41189', '412', 
              '4130', '4131', '4139', 
              '41400', '41401', '41402', 
              '41403', '41404', '41405', '41406', '41407', 
              '41410', '41411', '41412', '41419', 
              '4142', '4143', '4148', '4149')
is_heart_icd9_prcdr <- c('0066', '3601', '3602', '3603', 
                         '3604', '3605', '3606', '3607', 
                         '3609', '3610', '3611', '3612', 
                         '3613', '3614', '3615', '3616', 
                         '3617', '3619', '362', '3631', '3632')
is_heart_hcpcs <- c('33510', '33511', '33512', '33513', '33514', '33515', '33516', '33517', '33518', 
                    '33519', '33521', '33522', '33523', '33533', '33534', '33535', '33536', '33542', 
                    '33545', '33548', '92975', '92977', '92980', '92982', '92995', '33140', '33141')
ihd_claims1 <- cohort_post_procedure %>% 
  filter(if_any(starts_with("ICD9_DGNS"), ~ . %in% is_heart_icd9_dgns))
ihd_claims2 <- cohort_post_procedure %>% 
  filter(if_any(starts_with("ICD9_PRCDR"), ~ . %in% is_heart_icd9_prcdr))
ihd_claims3 <- cohort_post_procedure %>% 
  filter(if_any(starts_with("HCPCS"), ~ . %in% is_heart_hcpcs))
ihd_claims <- bind_rows(ihd_claims1, ihd_claims2, ihd_claims3) %>% 
  mutate(outcome_type = "ihd")

# Find any claims for cardiac arrest
ca_claims <- cohort_post_procedure %>% 
  filter(if_any(starts_with("ICD9_DGNS"), ~ str_starts(., pattern = "4275"))) %>% 
  mutate(outcome_type = "ca")

# Combine all cardiac events
cardiac_events <- bind_rows(ca_claims, ihd_claims, hf_claims, ang_claims)

# Check counts per procedure type
cardiac_events %>% 
  distinct(ID, procedure) %>% 
  group_by(procedure) %>% 
  count()
  

```

## Question: Why could this be other than the procedure itself?

# Filter the claims for comorbidities
### hypertension, high cholesterol, diabetes, and previous ischemic heart disease

```{r}

# Find any claims for hypertension
hyp_claims1 <- cohort_comorbidities %>% 
  filter(if_any(starts_with("ICD9_DGNS"), ~ str_starts(., pattern = "4011")))
hyp_claims2 <- cohort_comorbidities %>% 
  filter(if_any(starts_with("ICD9_DGNS"), ~ str_starts(., pattern = "4019")))
hyp_claims3 <- cohort_comorbidities %>% 
  filter(if_any(starts_with("ICD9_DGNS"), ~ str_starts(., pattern = "4010")))
hyp_claims <- bind_rows(hyp_claims1, hyp_claims2, hyp_claims3) %>% 
  mutate(comorbid_type = "hyp")

# Find any claims for high cholesterol
hchol_claims <- cohort_comorbidities %>% 
  filter(if_any(starts_with("ICD9_DGNS"), ~ str_starts(., pattern = "2720"))) %>% 
  mutate(comorbid_type = "hchol")

# Find any claims for diabetes
dia_icd9 <- c('249', '250', '3572', '36201', '36202', '36641')
dia_claims <- cohort_comorbidities %>% 
  filter(if_any(starts_with("ICD9_DGNS"), ~ . %in% dia_icd9)) %>% 
   mutate(comorbid_type = "dia")

# Find any claims for ischemic heart disease
ihd_claims1_co <- cohort_comorbidities %>% 
  filter(if_any(starts_with("ICD9_DGNS"), ~ . %in% is_heart_icd9_dgns))
ihd_claims2_co <- cohort_comorbidities %>% 
  filter(if_any(starts_with("ICD9_PRCDR"), ~ . %in% is_heart_icd9_prcdr))
ihd_claims3_co <- cohort_comorbidities %>% 
  filter(if_any(starts_with("HCPCS"), ~ . %in% is_heart_hcpcs))
ihd_claims_co <- bind_rows(ihd_claims1_co, ihd_claims2_co, ihd_claims3_co) %>% 
  mutate(comorbid_type = "ihd")

# Combine all comorbidities
comorbidities <- bind_rows(ihd_claims_co, dia_claims, hchol_claims, hyp_claims)

# Check counts per procedure type
comorbidities %>% 
  distinct(ID, procedure) %>% 
  group_by(procedure) %>% 
  count()

```

## Question: Why could this be other than the procedure itself?

## Potential next steps: 
Determine if the differences are significant: 
Build a logistic regression model on outcome using patient demographics and comorbidities as covariates

